---
title: "gmailr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{gmailr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(gmailr)
```

## Quickstart ##

This is a short document on getting up to speed with gmailr quickly.

## External Setup ##

In order to use gmailr you will need to create a google project for it. The
easiest way to do this is via the [Python
Quickstart](https://developers.google.com/gmail/api/quickstart/python).

* Click the `Enable the Gmail API` button.
* In the resulting dialog click the `DOWNLOAD CLIENT CONFIGURATION` on your computer.
* Tell gmailr where the JSON lives, by doing one of the two things
  1. Call `gm_auth_configure(path = "path/to/downloaded/json")
  2. Set the `GMAILR_APP` environment variable to the location of the JSON
     file, it is convienent to do this in your `.Renviron` file with
     `usethis::edit_r_environ()`. Then calling `gm_auth_configure()` with no arguments.
* Call `gm_auth()` to start the OAuth flow to verify to google that you would
  like your gmailr project to have access to your email. You will get a scary
  warning about an untrusted application, this is because the application is
  the one you just created, click advanced and `Go to gmailr` to proceed to do
  the oauth flow.
* If you want to authenticate with fewer scopes than the default use the
  `scopes` parameter to `gm_auth()`. You can see a full list of available
  scopes from `gm_scopes()`.
  
Only very heavy usage of the Gmail API requires payment, so use of the API for most
people should be free.

If you use `usethis::edit_r_environ()` to set both `GMAILR_EMAIL` and
`GMAILR_APP`, then once you have an oauth token you can simply run
`gm_auth_configure()` with no arguments at the top of the script to setup your
application.

## Writing new emails ##

Create a new email with `gm_mime()` and the helper functions. When testing it
is recommended to use `gm_create_draft()` to verify your email is formatted as you
expect before automating it (if desired) with `gm_send_message()`.

```r
test_email <-
  gm_mime() %>%
  gm_to("PUT_A_VALID_EMAIL_ADDRESS_THAT_YOU_CAN_CHECK_HERE") %>%
  gm_from("PUT_THE_GMAIL_ADDRESS_ASSOCIATED_WITH_YOUR_GOOGLE_ACCOUNT_HERE") %>%
  gm_subject("this is just a gmailr test") %>%
  gm_text_body("Can you hear me now?")

# Verify it looks correct
gm_create_draft(test_email)

# If all is good with your draft, then you can send it
gm_send_message(test_email)
```

You can add a file attachment to your message with `gm_attach_file()`.

```r
write.csv("mtcars.csv", mtcars)
test_email <- gm_attach_file("mtcars.csv")

# Verify it looks correct
gm_create_draft(test_email)

# If so, send it
gm_send_message(test_email)
```

## Reading emails ##

gmail shows you threads of messages in the web UI, you can retrieve all threads
with `gm_threads()`, and retrieve a specific thread with `gm_thread()`

```r
# view the latest thread
my_threads <- gm_threads(num_results = 10)

# retrieve the latest thread by retrieving the first ID

latest_thread <- gm_thread(gm_id(my_threads)[[1]])

# The messages in the thread will now be in a list
latest_thread$messages

# Retrieve parts of a specific message with the accessors
my_msg <- latest_thread$messages[[1]]

gm_to(my_msg)
gm_from(my_msg)
gm_date(my_msg)
gm_subject(my_msg)
gm_body(my_msg)

# If a message has attachments, download them all locally with `gm_save_attachments()`.
gm_save_attachments(my_msg)
```
