---
title: "Sending Messages With Gmailr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Sending Messages With Gmailr}
  %\usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
library(gmailr)
```

## Constructing a MIME message

### Text

First we will construct a simple text only message

```{r sending_messages_simple}
mime() %>%
  to("james.f.hester@gmail.com") %>%
  from("me@somewhere.com") %>%
  text_body("Gmailr is a very handy package!") -> text_msg
```

You can convert the message to a properly formatted MIME message using `as.character()`.

```{r sending_messages_simple_print}
strwrap(as.character(text_msg))
```

### HTML

You can also construct html messages.  It is customary to provide a text
only message along with the html message, but with modern email clients this is
not strictly necessary.

```{r sending_messages_html}
mime() %>%
  to("james.f.hester@gmail.com") %>%
  from("me@somewhere.com") %>%
  html_body("<b>Gmailr</b> is a <i>very</i> handy package!") -> html_msg
```

### Attachments

You can add attachments to your message in two ways.

1. If the data is in a file, use `attach_file()`.  The mime type is
   automatically guessed by `mime::guess_type`, or you can specify it yourself
   with the `type` parameter.
```{r sending_messages_attachments_2}
write.csv(file = "iris.csv", iris)

html_msg %>%
  subject("Here are some flowers") %>%
  attach_file("iris.csv") -> file_attachment
```

2. If the data are already loaded into R, you can use `attach_part()` to attach the binary data to your file.
```{r sending_messages_attachments_1}
html_msg %>% attach_part(part = charToRaw("attach me!"), name = "please") -> simple_attachment
```

## Uploading
### Create Draft

You can upload any mime message into your gmail drafts using `create_draft()`.
Be sure to give yourself at least `compose` permissions first.

```{r sending_messages_create_draft, eval=FALSE}
create_draft(file_attachment)
```

### Insert

This inserts the message directly into your mailbox, bypassing gmail's default
scanning and classification algorithms.

```{r sending_messages_insert_message, eval=FALSE}
insert_message(file_attachment)
```

### Import

This imports the email as though it was a normal message, with the same
scanning and classification as normal email.

```{r sending_messages_file_attachment, eval=FALSE}
insert_message(file_attachment)
```

## Sending

### Draft

`send_draft()` sends an email using the `draft_id` of an existing draft
(possibly created with `create_draft()`).

```{r sending_messages_send_draft, eval=FALSE}
my_drafts <- drafts()

send_draft(id(my_drafts, "draft_id")[1])
```
### Message

You can also send an email message directly from a `mime` object using `send_message()`.

```{r sending_messages_send_message, eval=FALSE}
send_message(file_attachment)
```

```{r sending_messages_clenup, include=FALSE}
unlink("iris.csv")
```
